---
title: "Exploring simulations"
author: "Max, Jim, Sean"
date: today
date-format: iso
toc: true
format: 
  html:
    page-layout: full
    embed-resources: true
knitr: 
  opts_chunk:
    fig.align: center
    out-width: 90%
editor: source
---

```{r load libraries}
#| message: false
#| warning: false

# Load libraries
library(tidyr)
library(dplyr)
library(patchwork)
library(RColorBrewer)
library(terra) # Load terra to use extract... needed for prep-data
library(ggtext)
library(ggstats)
library(ggdist)
library(ggplot2)
library(ggsidekick); theme_set(theme_sleek())
```


## Simulate data

### Experiment 1: cross-experiment with diffusion and null model for simulating data and retrieving parameters

- Simulate from yellowfin model
- Vary sigma and the diffusion parameters. How well can we retrieve them?

#### Fit a model to adult yellowfin sole
... To get a feeling for some sensible starting values for parameters and to get a model to simulate form

```{r load data}
#| warning: false
#| results: hide

source(here::here("analysis/prep-data.R"))
setwd(tmb_dir)
compile("movement_kernel.cpp", framework = "TMBad")
dyn.load(dynlib("movement_kernel"))
# dyn.unload(dynlib("movement_kernel"))
```

#### Fit a model to adult yellowfin sole

```{r}
species_set <- colnames(region_data_all)[30:ncol(region_data_all)]
N_c <- colSums(ifelse(region_data_all[, species_set] > 0, 1, 0))
species_set <- species_set[N_c > 1000]

param_set <- c("obj_diffusion", "obj_null", "deltaAIC", "range", "ln_kappa2", "corr_depth")

cI <- match("a_yfs", species_set)
species <- species_set[cI]

f_depth
if (f_depth == "identity") {
  depthprime_s <- (depth_s / 100)
  #depthprime_s <- (depth_s - mean(depth_s)) / sd(depth_s) # ML: Can we better estimate depth and depth sq if we square?
  #depthprime_s <- (log(depth_s) - mean(log(depth_s))) / sd(log(depth_s)) # ML: Can we better estimate depth and depth sq if we square?
} else if (f_depth == "log") {
  depthprime_s <- log(depth_s) - 4.5
}

# cor(depthprime_s, depthprime_s^2)
# depthprime_scaled <- (log(depth_s) - mean(log(depth_s))) / sd(log(depth_s)) # ML: Can we better estimate depth and depth sq if we square?
# cor(depthprime_scaled, depthprime_scaled^2)

Params <- list(
  "beta0" = 0,
  #"beta_j" = c(0.1, 0.1),
  "beta_j" = 0.1,
  "ln_tau" = 0,
  "ln_kappa" = 0,
  "omega_s" = rnorm(nrow(spde$c0)),
  "ln_kappa2" = exp(-1)
)

Random <- "omega_s"

# Special stuff
#if (Data$dist == "LNP") {
Params$ln_sigma_eta <- log(0.1)
Params$eta_i <- rnorm(nrow(region_data_all))
Random <- c(Random, "eta_i")
```

##### Diffusion model

```{r fit diffusion model to a_yfs}
Data_sim_diff <- list(
  "method" = "diffusion",
  "dist" = distribution,
  "c_i" = region_data_all[, species],
  "A_is" = A_is,
  "A_gs" = A_gs,
  "M0" = spde$c0,
  "M1" = spde$g1,
  "M2" = spde$g2,
  "invsqrtM0" = invsqrtM0,
  "invM0" = invM0,
  "depth_s" = depthprime_s,
  "sim_gmrf" = 0L #< new
)

Obj_sim_diff <- MakeADFun(
  data = Data_sim_diff,
  parameters = Params,
  # profile = c("beta0","beta_j"),    # ,"ln_kappa2"
  random = Random,
  checkParameterOrder = TRUE,
  DLL = "movement_kernel"
)

Obj_sim_diff$env$beSilent()

# Optimize
Opt_sim_diff <- nlminb(
  start = Obj_sim_diff$par,
  obj = Obj_sim_diff$fn,
  grad = Obj_sim_diff$gr,
  control = list(eval.max = 1e4, iter.max = 1e4, trace = 1)
)
```

##### Null model

```{r fit null model to a_yfs}
Data_sim_null <- list(
  "method" = "null",
  "dist" = distribution,
  "c_i" = region_data_all[, species],
  "A_is" = A_is,
  "A_gs" = A_gs,
  "M0" = spde$c0,
  "M1" = spde$g1,
  "M2" = spde$g2,
  "invsqrtM0" = invsqrtM0,
  "invM0" = invM0,
  "depth_s" = depthprime_s,
  "sim_gmrf" = 0L #< new
)

Obj_sim_null <- MakeADFun(
  data = Data_sim_null,
  parameters =  Params[setdiff(names(Params), c("ln_tau2","ln_kappa2"))],
  # profile = c("beta0","beta_j"),    # ,"ln_kappa2"
  random = Random,
  checkParameterOrder = TRUE,
  DLL = "movement_kernel"
)
Obj_sim_null$env$beSilent()

# Optimize
Opt_sim_null <- nlminb(
  start = Obj_sim_null$par,
  obj = Obj_sim_null$fn,
  grad = Obj_sim_null$gr,
  control = list(eval.max = 1e4, iter.max = 1e4, trace = 1)
)
```

#### Simulate `n_sim` data sets from both the diffussion operating model and the null model

```{r set fixed effects}
set.seed(99)

n_sim <- 100

# Diffusion operating model
p_sim_diff <- Obj_sim_diff$env$last.par.best

p_sim_diff[names(p_sim_diff) %in% c("beta0", "beta_j", "ln_sigma_eta",
                                    "ln_tau", "ln_kappa", "ln_kappa2")] 
  
# Replace fitted parameters with fake, more rounded values
p_sim_diff[names(p_sim_diff) == "beta0"] <- 2.9
#p_sim_diff[names(p_sim_diff) == "beta_j"] <- c(-5.5, 0.02)
p_sim_diff[names(p_sim_diff) == "beta_j"] <- -2.8
p_sim_diff[names(p_sim_diff) == "ln_tau"] <- -1.8
p_sim_diff[names(p_sim_diff) == "ln_kappa"] <- -0.8
p_sim_diff[names(p_sim_diff) == "ln_kappa2"] <- 1.4
p_sim_diff[names(p_sim_diff) == "ln_sigma_eta"] <- 0.25 

# Lower ln_sigma_eta
p_sim_diff2 <- p_sim_diff
p_sim_diff2[names(p_sim_diff) == "ln_sigma_eta"] <- 0.1

# Increase ln_sigma_eta
p_sim_diff3 <- p_sim_diff
p_sim_diff3[names(p_sim_diff3) == "ln_sigma_eta"] <- 0.5


# Null operating model
p_sim_null <- Obj_sim_null$env$last.par.best

p_sim_null[names(p_sim_null) %in% c("beta0", "beta_j", "ln_sigma_eta",
                                    "ln_tau", "ln_kappa", "ln_kappa2")] 
  
# Replace fitted parameters with fake, more rounded values
p_sim_null[names(p_sim_null) == "beta0"] <- 0.6
#p_sim_null[names(p_sim_null) == "beta_j"] <- c(-5.5, 0.02)
p_sim_null[names(p_sim_null) == "beta_j"] <- -1.2
# Keep these the same as for the diffusion model... 
p_sim_null[names(p_sim_null) == "ln_tau"] <- -1.8
p_sim_null[names(p_sim_null) == "ln_kappa"] <- -0.8
p_sim_null[names(p_sim_null) == "ln_kappa2"] <- 1.4
p_sim_null[names(p_sim_null) == "ln_sigma_eta"] <- 0.25 

# Lower ln_sigma_eta
p_sim_null2 <- p_sim_null
p_sim_null2[names(p_sim_null) == "ln_sigma_eta"] <- 0.1

# Increase ln_sigma_eta
p_sim_null3 <- p_sim_null
p_sim_null3[names(p_sim_null3) == "ln_sigma_eta"] <- 0.5


# Now simulate 3 data sets with different ln_sigma_eta for each operating model
sim_dat <- list()

for(i in 1:n_sim) {

  # Operating model = diffusion
  sim_025_diff <- tibble(c_i = Obj_sim_diff$simulate(par = p_sim_diff, complete = FALSE)$c_i) %>% 
    mutate(ln_sigma_eta = "0.25",
           operating_model = "diffusion",
           iter = i) 
  
  sim_01_diff <- tibble(c_i = Obj_sim_diff$simulate(par = p_sim_diff2, complete = FALSE)$c_i) %>%
    mutate(ln_sigma_eta = "0.1",
           operating_model = "diffusion",
           iter = i)
  
  sim_50_diff <- tibble(c_i = Obj_sim_diff$simulate(par = p_sim_diff3, complete = FALSE)$c_i) %>% 
    mutate(ln_sigma_eta = "0.5",
           operating_model = "diffusion",
           iter = i) 
  
  # Operating model = null
  sim_025_null <- tibble(c_i = Obj_sim_null$simulate(par = p_sim_null, complete = FALSE)$c_i) %>% 
    mutate(ln_sigma_eta = "0.25",
           operating_model = "null",
           iter = i) 
  
  sim_01_null <- tibble(c_i = Obj_sim_null$simulate(par = p_sim_null2, complete = FALSE)$c_i) %>%
    mutate(ln_sigma_eta = "0.1",
           operating_model = "null",
           iter = i)
  
  sim_50_null <- tibble(c_i = Obj_sim_null$simulate(par = p_sim_null3, complete = FALSE)$c_i) %>% 
    mutate(ln_sigma_eta = "0.5",
           operating_model = "null",
           iter = i) 

  
  
  sim_dat[[i]] <- bind_rows(
    sim_025_diff, sim_01_diff, sim_50_diff,
    sim_025_null, sim_01_null, sim_50_null
    )

}

sim_df <- bind_rows(sim_dat) %>% 
  mutate(index = paste(iter, ln_sigma_eta, sep = "-"))
```

#### Fit models to simulated data

I will fit two models; with and without diffusion. In this example, I draw 100 simulated data sets (based on the "rounded" fixed effect estimates while keeping the GMRF to the adult yellowfin sole values) in a for-loop and fit both models, then I extract the estimates.

```{r fit to simulated data}
#| results: hide
# These are the parameters I'm tracking
pars <- "beta_j" #c("beta_j", "beta_j")

# Storage lists for each operating and estimating model
pars_diff_list_sim_diff <- list()
pars_null_list_sim_diff <- list()
pars_diff_list_sim_null <- list()
pars_null_list_sim_null <- list()

# Loop through all n data sets 
for(i in unique(sim_df$index)) {
  
  # Operating model = diffusion
  Data_sim_diff <- Data_sim_diff
  
  Data_sim_diff$c_i <- sim_df %>%
    filter(index == i & operating_model == "diffusion") %>% 
    pull(c_i)

  # Diffusion model
  Data_sim_diff$method <- "diffusion"
  
  Obj_diff <- MakeADFun(
    data = Data_sim_diff,
    parameters = Params,
    random = Random,
    checkParameterOrder = TRUE,
    DLL = "movement_kernel"
    )
  
  Opt_diff <- nlminb(
    start = Obj_diff$par,
    obj = Obj_diff$fn,
    grad = Obj_diff$gr,
    control = list(eval.max = 1e4, iter.max = 1e4)
    )
  
  pars_diff_list_sim_diff[[i]] <- 
    tibble(par = c("beta[j1]"),
    #tibble(par = c("beta[j1]", "beta[j2]"),
           est = Opt_diff$par[names(Opt_diff$par) %in% pars]) %>%
    mutate(index = i, 
           `Estimation model` = "Diffusion",
           `Operating model` = "Operating model = Diffusion"
           )

  
  # Now do the model without diffusion
  Data_sim_diff$method <- "null"
  Params_null = Params[setdiff(names(Params), c("ln_tau2", "ln_kappa2"))]
  
  Obj_null <- MakeADFun(
    data = Data_sim_diff,
    parameters = Params_null,
    random = Random,
    checkParameterOrder = TRUE,
    DLL = "movement_kernel"
    )
  
  Opt_null <- nlminb(
    start = Obj_null$par,
    obj = Obj_null$fn,
    grad = Obj_null$gr,
    control = list(eval.max = 1e4, iter.max = 1e4)
    )
  
  pars_null_list_sim_diff[[i]] <-
    tibble(par = c("beta[j1]"),
    #tibble(par = c("beta[j1]", "beta[j2]"),
           est = Opt_null$par[names(Opt_null$par) %in% pars]) %>%
    mutate(index = i, 
           `Estimation model` = "Null",
           `Operating model` = "Operating model = Diffusion")
  
  
  
  
  
  
  # Operating model = null
  Data_sim_null <- Data_sim_null
  
  Data_sim_null$c_i <- sim_df %>%
    filter(index == i & operating_model == "null") %>% 
    pull(c_i)

  # Diffusion model
  Data_sim_null$method <- "diffusion"
  
  Obj_diff <- MakeADFun(
    data = Data_sim_null,
    parameters = Params,
    random = Random,
    checkParameterOrder = TRUE,
    DLL = "movement_kernel"
    )
  
  Opt_diff <- nlminb(
    start = Obj_diff$par,
    obj = Obj_diff$fn,
    grad = Obj_diff$gr,
    control = list(eval.max = 1e4, iter.max = 1e4)
    )
  
  pars_diff_list_sim_null[[i]] <- 
    tibble(par = c("beta[j1]"),
    #tibble(par = c("beta[j1]", "beta[j2]"),
           est = Opt_diff$par[names(Opt_diff$par) %in% pars]) %>%
    mutate(index = i, 
           `Estimation model` = "Diffusion",
           `Operating model` = "Operating model = Null"
           )

  
  # Now do the model without diffusion
  Data_sim_null$method <- "null"
  Params_null = Params[setdiff(names(Params), c("ln_tau2", "ln_kappa2"))]
  
  Obj_null <- MakeADFun(
    data = Data_sim_null,
    parameters = Params_null,
    random = Random,
    checkParameterOrder = TRUE,
    DLL = "movement_kernel"
    )
  
  Opt_null <- nlminb(
    start = Obj_null$par,
    obj = Obj_null$fn,
    grad = Obj_null$gr,
    control = list(eval.max = 1e4, iter.max = 1e4)
    )
  
  pars_null_list_sim_null[[i]] <-
    tibble(par = c("beta[j1]"),
    #tibble(par = c("beta[j1]", "beta[j2]"),
           est = Opt_null$par[names(Opt_null$par) %in% pars]) %>%
    mutate(index = i, 
           `Estimation model` = "Null",
           `Operating model` = "Operating model = Null")
  
}
```

```{r}
# Make a dataframe for plotting
pars_df <- bind_rows(
  bind_rows(pars_diff_list_sim_diff),
  bind_rows(pars_null_list_sim_diff),
  bind_rows(pars_diff_list_sim_null),
  bind_rows(pars_null_list_sim_null)
  ) %>% 
  separate_wider_delim(index, delim = "-", names = c("iter", "ln_sigma_eta"))
```

Plot how well the models can retrieve the true depth coefficients

```{r}
#| fig-height: 7
# True values (horizontal lines)
true <- tibble(value = c(as.numeric(p_sim_diff[names(p_sim_diff) == "beta_j"]),
                         as.numeric(p_sim_null[names(p_sim_null) == "beta_j"])),
               par = c("beta[j1]"),
               `Operating model` = c("Operating model = Diffusion",
                                     "Operating model = Null"), 
               )

pal <- colorRampPalette(brewer.pal(10, "BrBG"))(20)[c(13, 15, 18)]

ggplot(pars_df, aes(`Estimation model`, est, fill = ln_sigma_eta, color = ln_sigma_eta)) +
  geom_jitter(position = position_jitterdodge(dodge.width = 0.85, jitter.width = 0.1, jitter.height = 0),
              alpha = 0.4, shape = 21, color = "white", size = 2) +
  stat_halfeye(alpha = 0.5, color = NA, position = position_dodge(width = 0.85)) +
  stat_pointinterval(position = position_dodge(width = 0.85), color = "grey20") +
  labs(y = "Estimated depth coefficients", x = "Estimation model") +
  geom_hline(data = true, aes(yintercept = value),
             alpha = 0.6, linetype = 2) +
  facet_wrap(~`Operating model`, scales = "free", ncol = 2) +
  scale_color_manual(values = pal, name = "") +
  scale_fill_manual(values = pal, name = "ln_sigma_eta") +
  guides(color = "none", fill = guide_legend(position = "inside")) +
  theme(legend.position.inside = c(0.07, 0.9))

ggsave(paste0(here::here(), "/results/sims.pdf"), width = 20, height = 13, unit = "cm")
  
# Actually, make the x variable the estimation model and color in three categories the sigma... the repeat for the other operating model
# pars_df %>%
#   mutate(`Estimation model` = ifelse(`Estimation model` == "Diffusion",
#                                      "Estimation model = Diffusion",
#                                      "Estimation model = Null")) %>%
#   ggplot(aes(ln_sigma_eta, est, fill = ln_sigma_eta, color = ln_sigma_eta)) +
#   geom_jitter(position = position_jitterdodge(dodge.width = 1, jitter.width = 0.1, jitter.height = 0),
#               alpha = 0.3) +
#   stat_halfeye(alpha = 0.5, color = NA, position = position_dodge(width = 1)) +
#   geom_boxplot(width = 0.1, fill = NA, outlier.colour = NA, position = position_dodge(width = 1)) +
#   labs(y = "Estimated depth coefficients", x = "") +
#   geom_hline(data = true, aes(yintercept = value),
#              alpha = 0.6, linetype = 2) +
#   #facet_wrap(~par, scales = "free", ncol = 1, labeller = label_parsed) +
#   #facet_wrap(~`Operating model`, scales = "free", ncol = 2) +
#   facet_grid(`Operating model` ~ `Estimation model`, scales = "free") +
#   scale_color_manual(values = pal, name = "") +
#   scale_fill_manual(values = pal, name = "") +
#   guides(fill = "none", color = "none") +
#   theme(strip.text.y.right = element_text(angle = 270),
#         legend.position = "bottom")
```

### Experiment 3: Simulating only data, not GMRFs

- Change the operating model to a non-diffusion one (i.e., simulate data from model parameters that do not show diffusion)
- Compare the estimation models
- Is this interesting or more for completeness sake?

```{r}
# TODO
```

### Experiment 4: Add in GMRFs to the simulations

- Repeat at least experiment 1--2? Or make a new experiment?

```{r}
# TODO
```




