---
title: "Plot case study results"
author: "Max, Jim, Sean"
date: today
date-format: iso
toc: true
format: 
  html:
    page-layout: full
    embed-resources: true
knitr: 
  opts_chunk:
    fig.align: center
    out-width: 90%
editor: source
---

```{r load libraries}
#| message: false
#| warning: false

# Load libraries
library(tidyr)
library(dplyr)
library(ggtext)
library(ggstats)
library(patchwork)
library(ggplot2)
library(ggsidekick); theme_set(theme_sleek())
library(terra) # Load terra to use extract... needed for prep-data
```

### Explore results from the case study

First quickly look at some results from the case study. Mainly I'm interested in the results for adult yellowfin sole, since it's used in the simple-tmb.R example showcasing the simulation. I guess if we want to do a factorial simulation experiment with two estimation models (diffusion on/off) and two simulated datasets (diffusion on/off) (+ varying other things on top of this, e.g., error and GMRF), we want a species where diffusion is found and is supported.

```{r load cs results}
# Read data
res <- read.csv(paste0(here::here(), "/results/2024-04-04_identity_LNP/Results_cz.csv")) %>%
  mutate("Diffusion\nfavoured" = ifelse(deltaAIC > 0,
                                       "N",
                                       "Y"),
         abs_delta_aic = ifelse(abs(deltaAIC) > 2,
                                "<2",
                                ">2")) %>%
  mutate(X = ifelse(X == "a_yfs",
                    paste0("<span style=\"color: ", "tomato3", "\">", X, "</span>"),
                    X))
```

Plot $\Delta$AIC, correlation between depth and diffused depth, as well as the range estimate. The range of range estimates is pretty long so I log them for now!

```{r plot cs results}
#| fig-height: 7

p1 <- ggplot(res, aes(deltaAIC, reorder(X, desc(deltaAIC)),
                      color = `Diffusion\nfavoured`, shape = abs_delta_aic)) +
  geom_point(fill = NA) +
  labs(x = "ΔAIC", y = "Species") +
  guides(color = guide_legend(position = "inside"),
         shape = guide_legend(position = "inside")) + 
  theme(legend.position.inside = c(0.77, 0.85)) +
  scale_shape_manual(values = c(19, 21), name = "abs(ΔAIC)") +
  scale_color_brewer(palette = "Dark2", direction = -1)

p2 <- ggplot(res, aes(corr_depth, reorder(X, desc(deltaAIC)))) +
  geom_point() +
  labs(x = "Correlation between depth\nand diffused depth",
       y = "Species")

p3 <- ggplot(res, aes(log(range), reorder(X, desc(deltaAIC)))) +
  geom_point() +
  labs(y = "Species")

p1 + p2 + p3 +
  plot_layout(axes = "collect") +
  plot_annotation(tag_levels = "A") &
  geom_stripped_rows(aes(y = X), inherit.aes = FALSE) &
  theme(axis.text.y = ggtext::element_markdown())
```

Looks like in adult yellowfin sole (in red axis text), the diffusion model is parsimonious, the correlation between diffused depth is intermediate (low correlations are very clear in the map, showing as very smoothed/diffused), and the range is not extreme either. This makes it a great example. It also seems in about half the species, the diffusion model is not parsimonious, whereas for the other half, either the diffusion model is parsimonious or they are similar. There doesn't seem to be an relationship between the $\Delta$AIC and the correlation coefficient, nor the range.
