---
title: "Running simulations"
author: "Max, Jim, Sean"
date: today
date-format: iso
toc: true
format: 
  html:
    page-layout: full
    embed-resources: true
knitr: 
  opts_chunk:
    fig.align: center
    out-width: 90%
editor: source
---

```{r load libraries}
#| message: false
#| warning: false

# Load libraries
library(tidyr)
library(dplyr)
library(patchwork)
library(RColorBrewer)
library(readr)
library(terra) # Load terra to use extract... needed for prep-data
library(ggtext)
library(ggstats)
library(ggdist)
library(ggplot2)
library(egg)
library(ggsidekick); theme_set(theme_sleek())
```


## Simulate data

### Cross-experiment with diffusion and null model for simulating data and retrieving parameters

- Simulate from juvenile Pacific cod model
- Vary sigma and the diffusion parameters. How well can we retrieve them?
- How well can we get the correct model with AIC?

#### Fit a model to adult yellowfin sole
... To get a feeling for some sensible starting values for parameters and to get a model to simulate form

```{r load data}
#| warning: false
#| results: hide

source(here::here("analysis/prep-ebs-data.R"))
setwd(tmb_dir)
compile("movement_kernel_sim.cpp", framework = "TMBad")
dyn.load(dynlib("movement_kernel_sim"))
# dyn.unload(dynlib("movement_kernel"))
```

```{r}
species_set <- colnames(region_data_all)[30:ncol(region_data_all)]
N_c <- colSums(ifelse(region_data_all[, species_set] > 0, 1, 0))
species_set <- species_set[N_c > 1000]

param_set <- c("obj_diffusion", "obj_null", "deltaAIC", "range", "ln_kappa2", "corr_depth")

cI <- match("j_pcod", species_set)
species <- species_set[cI]

f_depth
if (f_depth == "identity") {
  #depthprime_s <- (depth_s / 100)
  depthprime_s <- (depth_s - mean(depth_s)) / sd(depth_s)
  #depthprime_s <- (log(depth_s) - mean(log(depth_s))) / sd(log(depth_s)) # ML: Can we better estimate depth and depth sq if we square?
} else if (f_depth == "log") {
  depthprime_s <- log(depth_s) - 4.5
}

# cor(depthprime_s, depthprime_s^2)
# depthprime_scaled <- (log(depth_s) - mean(log(depth_s))) / sd(log(depth_s)) # ML: Can we better estimate depth and depth sq if we square?
# cor(depthprime_scaled, depthprime_scaled^2)

Params <- list(
  "beta0" = 0,
  #"beta_j" = c(0.1, 0.1),
  "beta_j" = 0.1,
  "ln_tau" = 0,
  "ln_kappa" = 0,
  "omega_s" = rnorm(nrow(spde$c0)),
  "ln_kappa2" = exp(-1)
)

Random <- "omega_s"

# Special stuff
#if (Data$dist == "LNP") {
Params$ln_sigma_eta <- log(0.1)
Params$eta_i <- rnorm(nrow(region_data_all))
Random <- c(Random, "eta_i")
```

##### Diffusion estimation model

```{r fit diffusion model to a_yfs}
#| results: hide
Data_sim_diff <- list(
  "method" = "diffusion",
  "dist" = distribution,
  "c_i" = region_data_all[, species],
  "A_is" = A_is,
  "A_gs" = A_gs,
  "M0" = spde$c0,
  "M1" = spde$g1,
  "M2" = spde$g2,
  "invsqrtM0" = invsqrtM0,
  "invM0" = invM0,
  "depth_s" = depthprime_s,
  "sim_gmrf" = 1L #< new
)

Obj_sim_diff <- MakeADFun(
  data = Data_sim_diff,
  parameters = Params,
  # profile = c("beta0","beta_j"),    # ,"ln_kappa2"
  random = Random,
  checkParameterOrder = TRUE,
  DLL = "movement_kernel_sim"
)

Obj_sim_diff$env$beSilent()

# Optimize
Opt_sim_diff <- nlminb(
  start = Obj_sim_diff$par,
  obj = Obj_sim_diff$fn,
  grad = Obj_sim_diff$gr,
  control = list(eval.max = 1e4, iter.max = 1e4, trace = 1)
)
```

##### Null estimation model

```{r fit null model to a_yfs}
#| results: hide

Data_sim_null <- list(
  "method" = "null",
  "dist" = distribution,
  "c_i" = region_data_all[, species],
  "A_is" = A_is,
  "A_gs" = A_gs,
  "M0" = spde$c0,
  "M1" = spde$g1,
  "M2" = spde$g2,
  "invsqrtM0" = invsqrtM0,
  "invM0" = invM0,
  "depth_s" = depthprime_s,
  "sim_gmrf" = 1L #< new
)

Obj_sim_null <- MakeADFun(
  data = Data_sim_null,
  parameters =  Params[setdiff(names(Params), c("ln_tau2","ln_kappa2"))],
  # profile = c("beta0","beta_j"),    # ,"ln_kappa2"
  random = Random,
  checkParameterOrder = TRUE,
  DLL = "movement_kernel_sim"
)
Obj_sim_null$env$beSilent()

# Optimize
Opt_sim_null <- nlminb(
  start = Obj_sim_null$par,
  obj = Obj_sim_null$fn,
  grad = Obj_sim_null$gr,
  control = list(eval.max = 1e4, iter.max = 1e4, trace = 1)
)
```

#### Simulate `n_sim` data sets from both the diffussion operating model and the null model

Here we in addition simulate some datasets with different values for ln_kappa2, which determine the level of smoothing. This is only relevant for the diffusion operating model.

First set some parameters based on the fits above

```{r set fixed effects}
# Diffusion operating model
p_sim_diff <- Obj_sim_diff$env$last.par.best

p_sim_diff[names(p_sim_diff) %in% c("beta0", "beta_j", "ln_sigma_eta",
                                    "ln_tau", "ln_kappa", "ln_kappa2")] 
  
# Replace fitted parameters with fake, more rounded values
p_sim_diff[names(p_sim_diff) == "beta0"] <- -0.4
p_sim_diff[names(p_sim_diff) == "beta_j"] <- -2.4
p_sim_diff[names(p_sim_diff) == "ln_tau"] <- -1.4
p_sim_diff[names(p_sim_diff) == "ln_kappa"] <- -0.8
p_sim_diff[names(p_sim_diff) == "ln_kappa2"] <- 2.5
p_sim_diff[names(p_sim_diff) == "ln_sigma_eta"] <- log(1)

# Lower ln_sigma_eta
p_sim_diff2 <- p_sim_diff
p_sim_diff2[names(p_sim_diff) == "ln_sigma_eta"] <- log(0.1)

# Increase ln_sigma_eta
p_sim_diff3 <- p_sim_diff
p_sim_diff3[names(p_sim_diff3) == "ln_sigma_eta"] <- log(2)

# Decrease ln_kappa2
p_sim_diff4 <- p_sim_diff
p_sim_diff4[names(p_sim_diff3) == "ln_kappa2"] <- 0

# Increase ln_kappa2 more
p_sim_diff5 <- p_sim_diff
p_sim_diff5[names(p_sim_diff5) == "ln_kappa2"] <- 5



# Null operating model
p_sim_null <- Obj_sim_null$env$last.par.best

p_sim_null[names(p_sim_null) %in% c("beta0", "beta_j", "ln_sigma_eta",
                                    "ln_tau", "ln_kappa", "ln_kappa2")] 
  
# Replace fitted parameters with fake, more rounded values
p_sim_null[names(p_sim_null) == "beta0"] <- -0.4
p_sim_null[names(p_sim_null) == "beta_j"] <- -2.4
p_sim_null[names(p_sim_null) == "ln_tau"] <- -1.4
p_sim_null[names(p_sim_null) == "ln_kappa"] <- -0.8

# Keep these same as diffusion OM
p_sim_null[names(p_sim_null) == "ln_sigma_eta"] <- log(1)

# Lower ln_sigma_eta
p_sim_null2 <- p_sim_null
p_sim_null2[names(p_sim_null) == "ln_sigma_eta"] <- log(0.1)

# Increase ln_sigma_eta
p_sim_null3 <- p_sim_null
p_sim_null3[names(p_sim_null3) == "ln_sigma_eta"] <- log(2)
```

Now loop through each simulation index and simulate data given different parameters and operating models.

```{r simulate data}
set.seed(99)
n_sim <- 200

sim_dat <- function(obj = obj, par = par, operating_model = NA,
                    ln_sigma_eta = NA, ln_kappa_2 = NA, iter = NA) {
  
  if( length(p_sim_null3[names(par) == "ln_kappa2"]) > 0 ) {
    
    ln_kappa2 <- as.numeric(par[names(par) == "ln_kappa2"])
    
  } else{ ln_kappa2 = NA }
  
  tibble(c_i = obj$simulate(par = par, complete = FALSE)$c_i) |> 
    mutate(ln_sigma_eta = as.numeric(par[names(par) == "ln_sigma_eta"]),
           ln_kappa2 = ln_kappa2,
           iter = iter, 
           operating_model = operating_model) 
  
}

sim_dat_list <- list()

for(i in 1:n_sim) {

  # Operating model = diffusion
  sim_1_diff <- sim_dat(obj = Obj_sim_diff, par = p_sim_diff, operating_model = "diffusion", iter = i)
  sim_01_diff <- sim_dat(obj = Obj_sim_diff, par = p_sim_diff2, operating_model = "diffusion", iter = i)
  sim_2_diff <- sim_dat(obj = Obj_sim_diff, par = p_sim_diff3, operating_model = "diffusion", iter = i)
  
  # In the operating model = diffusion, we also simulate with ln_kappa_2, the smoothness of the diffusion for ln_sigma_eta = 0.25
  sim_1_diff_kappa2_neg2  <- sim_dat(obj = Obj_sim_diff, par = p_sim_diff4, operating_model = "diffusion", iter = i)
  sim_1_diff_kappa2_6 <- sim_dat(obj = Obj_sim_diff, par = p_sim_diff5, operating_model = "diffusion",iter = i)
  
  # Operating model = null
  sim_1_null <- sim_dat(obj = Obj_sim_null, par = p_sim_null, operating_model = "null", iter = i)
  sim_01_null <- sim_dat(obj = Obj_sim_null, par = p_sim_null2, operating_model = "null", iter = i)
  sim_2_null <- sim_dat(obj = Obj_sim_null, par = p_sim_null3, operating_model = "null", iter = i)

  sim_dat_list[[i]] <- bind_rows(
    sim_1_diff, sim_01_diff, sim_2_diff,
    sim_1_diff_kappa2_neg2, sim_1_diff_kappa2_6,
    sim_1_null, sim_01_null, sim_2_null
    ) |> 
    mutate(index = paste(iter, ln_sigma_eta, sep = ","))

}

sim_df <- bind_rows(sim_dat_list)
```

#### Fit models to simulated data

I will fit two models; with and without diffusion. In this example, I draw 100 simulated data sets (based on the "rounded" fixed effect estimates while keeping the GMRF to the adult yellowfin sole values) in a for-loop and fit both models, then I extract the estimates.

```{r fit to simulated data}
#| results: hide

# These are the parameters I'm tracking
pars <- "beta_j"

# Storage lists for each operating and estimating model
pars_diff_list_sim_diff <- list()
pars_null_list_sim_diff <- list()
pars_diff_list_sim_null <- list()
pars_null_list_sim_null <- list()
aic_list_diff_om <- list()
aic_list_null_om <- list()

# In this loop, we want to keep ln_kappa2 as NA (null operating model) AND ln_kappa2 = median,
# because we simulated more data for the diffusion operating model since we also varied ln_kappa2
sim_df_main <- sim_df |>
  mutate(keep = replace_na(ln_kappa2, median(ln_kappa2, na.rm = TRUE))) |> 
  filter(keep == median(ln_kappa2, na.rm = TRUE))

# Loop through all n data sets
for(i in unique(sim_df_main$index)) {

  # Operating model = diffusion
  Data_sim_diff_i <- Data_sim_diff

  Data_sim_diff_i$c_i <- sim_df_main |>
    filter(index == i & operating_model == "diffusion") |>
    pull(c_i)

  # EM: Diffusion
  Data_sim_diff_i$method <- "diffusion"

  Obj_diff <- MakeADFun(
    data = Data_sim_diff_i,
    parameters = Params,
    random = Random,
    checkParameterOrder = TRUE,
    DLL = "movement_kernel_sim"
    )

  Opt_diff <- nlminb(
    start = Obj_diff$par,
    obj = Obj_diff$fn,
    grad = Obj_diff$gr,
    control = list(eval.max = 1e4, iter.max = 1e4)
    )

  pars_diff_list_sim_diff[[i]] <-
    tibble(par = c("beta[j1]"),
           est = Opt_diff$par[names(Opt_diff$par) %in% pars]) |>
    mutate(index = i,
           `Estimation model` = "Diffusion",
           `Operating model` = "Operating model = Diffusion"
           )


  # EM: Null
  Data_sim_diff_i$method <- "null"
  Params_null = Params[setdiff(names(Params), c("ln_tau2", "ln_kappa2"))]

  Obj_null <- MakeADFun(
    data = Data_sim_diff_i,
    parameters = Params_null,
    random = Random,
    checkParameterOrder = TRUE,
    DLL = "movement_kernel_sim"
    )

  Opt_null <- nlminb(
    start = Obj_null$par,
    obj = Obj_null$fn,
    grad = Obj_null$gr,
    control = list(eval.max = 1e4, iter.max = 1e4)
    )

  pars_null_list_sim_diff[[i]] <-
    tibble(par = c("beta[j1]"),
           est = Opt_null$par[names(Opt_null$par) %in% pars]) |>
    mutate(index = i,
           `Estimation model` = "Null",
           `Operating model` = "Operating model = Diffusion")

  # Delta AIC for the two EMs and Diffusion OM
  # if null is larger than diff, diff is favoured, and aic is [1, 0], else its [0, 1]
  if( (2*Opt_null$objective+2*length(Opt_null$par)) -
      (2*Opt_diff$objective+2*length(Opt_diff$par)) > 0 ){
    aic = c(1, 0)
  } else{aic = c(0, 1)}

  aic_list_diff_om[[i]] <-
    tibble(`Operating model` = "Diffusion",
           `Estimation model` = c("Diffusion", "Null"),
           aic = aic
           ) |>
    mutate(index = i)




  # Operating model = null
  # Getting lots of errors like these: inner newton optimization failed during gradient calculation
  Data_sim_null_i <- Data_sim_null

  Data_sim_null_i$c_i <- sim_df_main |>
    filter(index == i & operating_model == "null") |>
    pull(c_i)

  # EM: Diffusion
  Data_sim_null_i$method <- "diffusion"

  Obj_diff <- MakeADFun(
    data = Data_sim_null_i,
    parameters = Params,
    random = Random,
    checkParameterOrder = TRUE,
    DLL = "movement_kernel_sim"
    )

  Opt_diff <- nlminb(
    start = Obj_diff$par,
    obj = Obj_diff$fn,
    grad = Obj_diff$gr,
    control = list(eval.max = 1e4, iter.max = 1e4)
    )

  pars_diff_list_sim_null[[i]] <-
    tibble(par = c("beta[j1]"),
           est = Opt_diff$par[names(Opt_diff$par) %in% pars]) |>
    mutate(index = i,
           `Estimation model` = "Diffusion",
           `Operating model` = "Operating model = Null"
           )


  # EM: null
  Data_sim_null_i$method <- "null"
  Params_null = Params[setdiff(names(Params), c("ln_tau2", "ln_kappa2"))]

  Obj_null <- MakeADFun(
    data = Data_sim_null_i,
    parameters = Params_null,
    random = Random,
    checkParameterOrder = TRUE,
    DLL = "movement_kernel_sim"
    )

  Opt_null <- nlminb(
    start = Obj_null$par,
    obj = Obj_null$fn,
    grad = Obj_null$gr,
    control = list(eval.max = 1e4, iter.max = 1e4)
    )

  pars_null_list_sim_null[[i]] <-
    tibble(par = c("beta[j1]"),
           est = Opt_null$par[names(Opt_null$par) %in% pars]) |>
    mutate(index = i,
           `Estimation model` = "Null",
           `Operating model` = "Operating model = Null")

  # Delta AIC for the two EMs and null OM
  if( (2*Opt_null$objective+2*length(Opt_null$par)) -
      (2*Opt_diff$objective+2*length(Opt_diff$par)) > 0 ){
    aic = c(1, 0)
  } else{aic = c(0, 1)}

  aic_list_null_om[[i]] <-
    tibble(`Operating model` = "Null",
           `Estimation model` = c("Diffusion", "Null"),
           aic = aic
           ) |>
    mutate(index = i)

}
```

```{r fit to simulated data vary ln_kappa2}
#| results: hide

# This only uses the diffusion OM 
# I track beta and kappa2
# I do not do the AIC confusion matrix since we don't look at accuracy of finding the right model, just parameter estimation
pars <- c("beta_j", "ln_kappa2")

# Storage lists for each operating and estimating model
pars_diff_list_sim_diff_kappa <- list()
pars_null_list_sim_diff_kappa <- list()
aic_list_diff <- list()

# In this loop, we want to keep ln_kappa2 as NA (null operating model) and ln_kappa2 = midpoint, because we simulated more data for the diffusion operating model since we also varied ln_kappa2
sim_df_kappa <- sim_df |> 
  drop_na(ln_kappa2) |> # this removes the simulations using the null OM
  mutate(index_k = paste(index, ln_kappa2, sep = ","))

# Loop through all n data sets 
for(i in unique(sim_df_kappa$index_k)) {
  
  # Operating model = diffusion
  Data_sim_diff_i <- Data_sim_diff
  
  Data_sim_diff_i$c_i <- sim_df_kappa |>
    filter(index_k == i & operating_model == "diffusion") |> 
    pull(c_i)

  # EM: Diffusion
  Data_sim_diff_i$method <- "diffusion"
  
  Obj_diff <- MakeADFun(
    data = Data_sim_diff_i,
    parameters = Params,
    random = Random,
    checkParameterOrder = TRUE,
    DLL = "movement_kernel_sim"
    )
  
  Opt_diff <- nlminb(
    start = Obj_diff$par,
    obj = Obj_diff$fn,
    grad = Obj_diff$gr,
    control = list(eval.max = 1e4, iter.max = 1e4)
    )
  
  pars_diff_list_sim_diff_kappa[[i]] <- 
    tibble(par = c("beta[j1]", "ln_kappa2"),
           est = Opt_diff$par[names(Opt_diff$par) %in% pars]) |>
    mutate(index = i, 
           `Estimation model` = "Diffusion",
           `Operating model` = "Operating model = Diffusion"
           )

  
  # EM: Null
  Data_sim_diff_i$method <- "null"
  Params_null = Params[setdiff(names(Params), c("ln_tau2", "ln_kappa2"))]
  
  Obj_null <- MakeADFun(
    data = Data_sim_diff_i,
    parameters = Params_null,
    random = Random,
    checkParameterOrder = TRUE,
    DLL = "movement_kernel_sim"
    )
  
  Opt_null <- nlminb(
    start = Obj_null$par,
    obj = Obj_null$fn,
    grad = Obj_null$gr,
    control = list(eval.max = 1e4, iter.max = 1e4)
    )
  
  pars_null_list_sim_diff_kappa[[i]] <-
    tibble(par = "beta[j1]",
           est = Opt_null$par[names(Opt_null$par) == "beta_j"]) |> # Do not track ln_kappa2 for the Null EM
    mutate(index = i, 
           `Estimation model` = "Null",
           `Operating model` = "Operating model = Diffusion")
  
  # Delta AIC for the two EMs and Diffusion OM
  # if null is larger than diff, diff is favoured, and aic is [1, 0], else its [0, 1]
  if( (2*Opt_null$objective+2*length(Opt_null$par)) -
      (2*Opt_diff$objective+2*length(Opt_diff$par)) > 0 ){
    aic = c(1, 0)
  } else{aic = c(0, 1)}
  
  aic_list_diff[[i]] <-
    tibble(`Operating model` = "Diffusion",
           `Estimation model` = c("Diffusion", "Null"),
           aic = aic
           ) |> 
    mutate(index = i)
    
}
```

```{r}
# Make dataframes for plotting
pars_df <- bind_rows(
  bind_rows(pars_diff_list_sim_diff),
  bind_rows(pars_null_list_sim_diff),
  bind_rows(pars_diff_list_sim_null),
  bind_rows(pars_null_list_sim_null)
  ) |>
  separate_wider_delim(index, delim = ",", names = c("iter", "ln_sigma_eta"))

write_csv(pars_df, paste0(here::here(), "/results/simulation/pars_df.csv"))
#pars_df <- read_csv(paste0(here::here(), "/results/simulation/pars_df.csv"))

pars_df_kappa <- bind_rows(
  bind_rows(pars_null_list_sim_diff_kappa),
  bind_rows(pars_diff_list_sim_diff_kappa)
  ) |> 
  separate_wider_delim(index, delim = ",", names = c("iter", "ln_sigma_eta", "ln_kappa2"))

write_csv(pars_df_kappa, paste0(here::here(), "/results/simulation/pars_df_kappa.csv"))
#pars_df_kappa <- read_csv(paste0(here::here(), "/results/simulation/pars_df_kappa.csv"))

aic_df <- bind_rows(
  bind_rows(aic_list_null_om),
  bind_rows(aic_list_diff_om)
  )

write_csv(aic_df, paste0(here::here(), "/results/simulation/aic_df.csv"))
#aic_df <- read_csv(paste0(here::here(), "/results/simulation/aic_df.csv"))

aic_df_kappa <- bind_rows(
  bind_rows(aic_list_diff)
  )

write_csv(aic_df_kappa, paste0(here::here(), "/results/simulation/aic_df_kappa.csv"))
#aic_df_kappa <- read_csv(paste0(here::here(), "/results/simulation/aic_df_kappa.csv"))
```

AIC confusion matrix

```{r}
aic_df |>
  separate(index, ",", into = c("index", "ln_sigma_eta")) |>
  mutate(sigma_eta = round(exp(as.numeric(ln_sigma_eta))), digits = 2) |> 
  summarise(count = sum(aic),
            .by = c(`Operating model`, `Estimation model`, sigma_eta)) |>
  ggplot(aes(`Operating model`, `Estimation model`, label = count)) +
  geom_text() +
  facet_wrap(~sigma_eta) +
  theme(aspect.ratio = 1)

ggsave(paste0(here::here(), "/results/simulation/confusion_table.pdf"), width = 15, height = 7, unit = "cm")
```

How does AIC difference between the estimation models and diffusion operating model change with ln_kappa?

```{r}
aic_df_kappa |> 
  separate_wider_delim(index, delim = ",", names = c("iter", "ln_sigma_eta", "ln_kappa2")) |> 
  filter(ln_sigma_eta == median(as.numeric(ln_sigma_eta))) |> 
  summarise(count = sum(aic),
            .by = c(`Estimation model`, ln_kappa2)) |> 
  ggplot(aes(ln_kappa2, `Estimation model`, label = count)) +
  geom_text() + 
  ggtitle("Operating model = Diffusion")

ggsave(paste0(here::here(), "/results/simulation/confusion_table_kappa.pdf"), width = 11, height = 11, unit = "cm")
```

Plot how well the models can retrieve the true depth coefficients given variation in ln_sigma_eta

```{r}
#| fig-height: 7
# True values (horizontal lines)
true <- tibble(value = c(as.numeric(p_sim_diff[names(p_sim_diff) == "beta_j"]),
                         as.numeric(p_sim_null[names(p_sim_null) == "beta_j"])),
               par = c("beta[j1]"),
               `Operating model` = c("Operating model = Diffusion",
                                     "Operating model = Null"),
               )

pal <- colorRampPalette(brewer.pal(10, "BrBG"))(20)[c(13, 15, 18)]

# Add in the AIC confusion scores
aic_df_sum <- aic_df |> 
  separate(index, ",", into = c("index", "ln_sigma_eta")) |>
  summarise(count = sum(aic),
            .by = c(`Operating model`, `Estimation model`, ln_sigma_eta)) |>
  mutate(count_prop = count / n_sim) |> 
  mutate(`Operating model` = ifelse(`Operating model` == "Diffusion", 
                                     "Operating model = Diffusion", 
                                     "Operating model = Null")) |> 
  mutate(sigma_eta = 1, 
         sigma_eta = ifelse(ln_sigma_eta == min(ln_sigma_eta), 0.1, sigma_eta),
         sigma_eta = ifelse(ln_sigma_eta == max(ln_sigma_eta), 2, sigma_eta),
         sigma_eta = as.character(sigma_eta)) |> 
  left_join(pars_df |> 
              mutate(ln_sigma_eta = as.character(ln_sigma_eta)) |> 
              summarise(est = max(est) + 0.1, .by = c(`Estimation model`, `Operating model`, ln_sigma_eta)))

pars_df <- pars_df |> 
  mutate(sigma_eta = 1, 
         sigma_eta = ifelse(ln_sigma_eta == min(ln_sigma_eta), 0.1, sigma_eta),
         sigma_eta = ifelse(ln_sigma_eta == max(ln_sigma_eta), 2, sigma_eta),
         sigma_eta = as.character(sigma_eta))

p <- ggplot(pars_df, aes(`Estimation model`, est, fill = sigma_eta, color = sigma_eta)) +
  geom_jitter(position = position_jitterdodge(dodge.width = 0.85, jitter.width = 0.1, jitter.height = 0),
              alpha = 0.3, shape = 21, color = "white", size = 2) +
  stat_halfeye(alpha = 0.5, color = NA, position = position_dodge(width = 0.85)) +
  stat_pointinterval(position = position_dodge(width = 0.85), color = "grey20") +
  labs(y = "Estimated depth coefficient", x = "Estimation model") +
  geom_hline(data = true, aes(yintercept = value),
             alpha = 0.6, linetype = 2) +
  labs(fill = c(expression(paste("Observation error SD, ", sigma[eta])))) +
  facet_wrap(~`Operating model`, ncol = 2, scales = "free") +
  scale_color_manual(values = pal, name = "") +
  scale_fill_manual(values = pal) +
  guides(color = "none",
         fill = guide_legend(#position = "inside",
                             title.position = "top", title.hjust = 0.5)) +
  geom_text(data = aic_df_sum, aes(label = round(count_prop, digits = 2), color = ln_sigma_eta),
            size = 3.5, position = position_dodge(width = 0.85), fontface = 2) +
  theme(#legend.position.inside = c(0.75, 0.15),
        legend.position = "bottom",
        legend.direction = "horizontal")

tag_facet(p, fontface = 1, size = 3.5)

ggsave(paste0(here::here(), "/results/figures/sims_beta_sigma.pdf"), width = 20, height = 13, unit = "cm")
```

Plot how well the models can retrieve the true depth coefficients (and kappa) with kappa variation and only the diffusion OM

```{r}
# True values (horizontal lines)
true <- tibble(value = as.numeric(p_sim_diff[names(p_sim_diff) == "beta_j"]),
               par = c("beta[j1]"))

# TODO: perhaps I wanted to swap so top row is " ... diffusion" and bottom row is kappa, but then I need to manually set the factor 
# levels in the fact, which messes up tag_facet... 

#pars_df_kappa |> distinct(ln_kappa2, ln_sigma_eta)

pars_df_kappa2 <- pars_df_kappa |>
  filter(par == "beta[j1]" & ln_sigma_eta == 0) |> 
  mutate(ln_kappa2 = ifelse(ln_kappa2 == 0, "log(\u03BA) = 0\nHigh diffusion", ln_kappa2),
         ln_kappa2 = ifelse(ln_kappa2 == 2.5, "log(\u03BA) = 2.5\nMedium diffusion", ln_kappa2),
         ln_kappa2 = ifelse(ln_kappa2 == 5, "log(\u03BA) = 5\nLow diffusion", ln_kappa2))

# Add in the AIC confusion scores
aic_df_kappa_sum <- aic_df_kappa |> 
  separate_wider_delim(index, delim = ",", names = c("iter", "ln_sigma_eta", "ln_kappa2")) |> 
  filter(ln_sigma_eta == 0) |> 
  summarise(count = sum(aic),
            .by = c(`Estimation model`, ln_kappa2)) |> 
  mutate(count_prop = round(count / n_sim, digits = 2)) |> 
  mutate(ln_kappa2 = ifelse(ln_kappa2 == 0, "log(\u03BA) = 0\nHigh diffusion", ln_kappa2),
         ln_kappa2 = ifelse(ln_kappa2 == 2.5, "log(\u03BA) = 2.5\nMedium diffusion", ln_kappa2),
         ln_kappa2 = ifelse(ln_kappa2 == 5, "log(\u03BA) = 5\nLow diffusion", ln_kappa2)) |> 
  left_join(pars_df_kappa2 |> 
              summarise(est = max(est) + 0.2, .by = c(`Estimation model`, ln_kappa2)))

# Depth coefficients
p <- ggplot(pars_df_kappa2, aes(`Estimation model`, est)) +
  geom_jitter(alpha = 0.15, shape = 21, color = "white", fill = "grey10", size = 2, height = 0, width = 0.1) +
  stat_halfeye(alpha = 0.8, color = NA) +
  stat_pointinterval(color = "grey20") +
  labs(y = "Estimated depth coefficient", x = "Estimation model") +
  geom_hline(data = true, aes(yintercept = value),
             alpha = 0.6, linetype = 2) +
  geom_text(data = aic_df_kappa_sum, aes(label = count_prop), size = 3.5, color = "grey30") + 
  facet_wrap(~ln_kappa2, ncol = 3) +
  guides(color = "none", fill = guide_legend(position = "inside")) +
  theme(legend.position.inside = c(0.07, 0.93))

tag_facet(p, fontface = 1, size = 3.5)

ggsave(paste0(here::here(), "/results/figures/sims_beta_kappa.pdf"),
       width = 20, height = 11, unit = "cm", device = cairo_pdf)
```

```{r}
# FIXME
# How important is:

#iter: 1  Error in iterate(par) : 
#   Newton dropout because inner gradient had non-finite components.
# In addition: Warning message:
# In nlminb(start = Obj_diff$par, obj = Obj_diff$fn, grad = Obj_diff$gr,  :
#   NA/NaN function evaluation
```

